Are kelp forests different? A comparison of kelp forests between contrasting study sites

### 2018-07-19 paper_draft.Rmd
references edited...more to come

---
title: "Are kelp forests different? A comparison of kelp forests between contrasting study sites"
output:
  pdf_document:
    toc: yes
  html_notebook:
    css: C:/Program Files/R/r_working_directory/benthic_paper/ecography.csl
    fig_caption: yes
    fig_height: 5
    fig_width: 4
    toc: yes
  word_document:
    toc: yes
---

@@

Kelps are a cold-water species of large brown algae, belonging to the order Laminariales [@Kelly 2005]. They are regarded as ecosystem engineers, providing a 3-dimensional habitat for a variety of fauna and flora. In addition, kelp ecosystems host a number of economically important species such as lobster, abalone and fish [@Dayton1985; @Steneck2002]. Various abiotic (e.g. light, wave exposure, sedimentation) and biotic (e.g. grazing) factors structure kelp forests [@Devinny1978; @Dayton1992; @Utter1996; @Toohey2004; @Rothman2006; @Sjøtun2006; @Wernberg2008; @Pedersen2012]. Variations in these factors ultimately affects the way kelp-associated benthos communities are structured [@Dayton1985; @Guimaraes1996; @Wernberg2008; @Wernberg 2009]
.
Temperature is recognised as the main abiotic driver of marine species distributions [@Dayton1992; @Hiscock2004; @Wernberg2011; @Wernberg2013; @Wernberg2016; @Mabin2013]. Given kelp’s preference for cold-temperate waters, ocean warming poses a serious threat to these ecosystems [@Wernberg2010; @Wernberg2011; @Verges2016]. In the absence of kelp, other types of systems may persist, for example algal turf, coralline reefs or urchin barrens [Verges2016]. These other systems support a different array of fauna and flora and are considered less diverse and less productive than kelp forests [@Dayton1985; @Connell2008; @Byrnes2011]. In future climate scenarios, species distributions are expected to shift poleward in response to changing ocean temperatures. Given kelps are cold-water species, changes in ocean temperatures, particularly warming, could have detrimental effects on their growth, survival, reproduction and recruitment.

Other abiotic drivers such as light [@Dayton1984; @Bolton1985; @Dieck1993; @Sjøtun2006; @Wernberg2008], sedimentation [@Duggins1990; @Vadas1992; @Schiel2006; Schiel2006] and water motion [@Hurd2000; @Kawamata2001; @Mudge2003; @Scrosati2004; @Wernberg2008; @Pedersen2012] have been shown to be important in the settlement, growth, and survival of kelp and biotic variables (e.g. see @Black1976; @Kennelly1983; @Dayton1984; @Harris1984; @Carpenter1990; @Sivertsen2006; @Harley2012) play an equally important role in structuring kelp communities, through alteration in top-down and bottom-up control mechanisms [@Breen1976; @Duggins1980; @Harris1984; @Estes1995; @Shears2002; @Arkema2009; @Blamey2012].

Kelp forest structure can be influenced by a variety of factors that ultimately determine the size frequency distribution of individuals and their spatially and temporarily variability. Much of the past research on kelp population structure focuses on the different variables that may affect kelp recruitment. It is generally accepted that most kelps rely on a ‘recruitment window’; a time when environmental factors are favourable for spore release and growth (Deysher and Dean 1986); however, other factors such as temperature, nutrients, light, wave motion, substratum, sedimentation, grazing etc. may also be important in the settlement and growth of spores, even during optimal recruiting conditions (Devinny and Volse 1978, Dayton 1985, Deysher and Dean 1986a, 1986b, Vadas et al. 1992, Kawamata 2001, Taylor and Schiel 2003, Schiel et al. 2006, Rothäusler et al. 2009, Staehr and Wernberg 2009, Wernberg et al. 2011, Gaitán-Espitia et al. 2014, Franco et al. 2015).

@@

Population structure of _E. maxima_ may depend on a variety of factors such as temperature [@Lee1988; @Raven1988; @Wiencke1989; @Dieck1993; @Connell2010; @Bartsch2012; @Mabin2013], light [@Lee1988; @Altamirano2004; @Wernberg2008; @Arkema2009; @Tait2014], wave motion [@Graham1997a; @Kawamata2001; @Wernberg2005; @FowlerWalker2006; @Pedersen2012] and nutrients [@Jackson1977; @AndrewsHutchings1980; @Field1980b; @Probyn1985; @Guimaraes1996; @Edwards2006]. At shallow depths (2–3m), the canopy tends to reach the surface; however, the structure changes to a mixture of sub-canopy and canopy at greater depths (5–10m), with eventual transition into _L. pallida_ dominated canopy (>10m) [@Rothman2006; @Rothman2010]. Therefore, from the surface it may appear that there is a change in biomass due to reduction in canopy cover with depth; however, due to the presence of sub-canopy kelps at deeper depths the biomass changes insignificantly [@Rothman2006]. Populations of E. maxima dominate the west coast, but more recently have increased in density in False Bay and further east [@Reimers2012], extending their range by 70km along the south coast (Bolton et al. 2012). The range extension of E. maxima has been associated with a possible cooling and increased nutrient supply in the nearshore around the South African south-west coast [@Moloney2005; @Blamey2015b].

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap= 'Multidimensional scaling of flora and fauna communities respectively.'}
ggarrange(nmds1_plot, nmds2_plot, ncol = 2, nrow = 1, labels = "AUTO", align = "none")
```

>>>>

---
title: "Are kelp forests different? A comparison of kelp forests between contrasting study sites"
output:
  pdf_document:
    toc: yes
  html_notebook:
    css: C:/Program Files/R/r_working_directory/benthic_paper/ecography.csl
    fig_caption: yes
    fig_height: 5
    fig_width: 4
    toc: yes
  word_document:
    toc: yes
bibliography: mycollection.bib
---

@@

Kelps are a cold-water species of large brown algae, belonging to the order Laminariales. They are regarded as ecosystem engineers, providing a 3-dimensional habitat for a variety of fauna and flora. In addition, kelp ecosystems host a number of economically important species such as lobster, abalone and fish [@Dayton1985; @Steneck2002]. Various abiotic (e.g. light, wave exposure, sedimentation) and biotic (e.g. grazing) factors structure kelp forests [@Devinny1978; @Dayton1992; @Utter1996; @Toohey2004; @Rothman2006; @Wernberg2008; @pedersen2012a]. Variations in these factors ultimately affects the way kelp-associated benthos communities are structured [@Dayton1985; @Guimaraes1996; @Wernberg2008; @Wernberg2009]
.
Temperature is recognised as the main abiotic driver of marine species distributions [@Dayton1992; @Hiscock2004; @Wernberg2011; @Wernberg2013; @Wernberg2016; @Mabin2013]. Given kelp’s preference for cold-temperate waters, ocean warming poses a serious threat to these ecosystems [@Wernberg2010; @Wernberg2011; @Verges2016]. In the absence of kelp, other types of systems may persist, for example algal turf, coralline reefs or urchin barrens [Verges2016]. These other systems support a different array of fauna and flora and are considered less diverse and less productive than kelp forests [@Dayton1985; @Connell2008; @Byrnes2011]. In future climate scenarios, species distributions are expected to shift poleward in response to changing ocean temperatures. Given kelps are cold-water species, changes in ocean temperatures, particularly warming, could have detrimental effects on their growth, survival, reproduction and recruitment.

Other abiotic drivers such as light [@Dayton1984; @Bolton1985a; @Dieck1993; @Sjøtun2006; @Wernberg2008], sedimentation [@Duggins1990; @Vadas1992; @Schiel2006; Schiel2006] and water motion [@Hurd2000; @Kawamata2001; @Mudge2003; @Scrosati2004; @Wernberg2008; @pedersen2012a] have been shown to be important in the settlement, growth, and survival of kelp and biotic variables (e.g. see @Black1976; @Kennelly1983; @Dayton1984; @Harris1984; @Carpenter1990; @Sivertsen2006; @Harley2012) play an equally important role in structuring kelp communities, through alteration in top-down and bottom-up control mechanisms [@Breen1976; @Duggins1980; @Harris1984; @Estes1995; @Shears2002; @Arkema2009; @Blamey2012].

Kelp forest structure can be influenced by a variety of factors that ultimately determine the size frequency distribution of individuals and their spatially and temporarily variability. Much of the past research on kelp population structure focuses on the different variables that may affect kelp recruitment. It is generally accepted that most kelps rely on a ‘recruitment window’; a time when environmental factors are favourable for spore release and growth [@Deysher1986]; however, other factors such as temperature, nutrients, light, wave motion, substratum, sedimentation, grazing etc. may also be important in the settlement and growth of spores, even during optimal recruiting conditions [@Devinny1978; @Dayton1985; @Deysher1986a; Deysher1986b; @Vadas1992; @Kawamata2001; @Taylor2003; @Schiel2006; @Rothausler2009; @Staehr2009; @Wernberg2011; @Franco2015].

@@

Population structure of _E. maxima_ may depend on a variety of factors such as temperature [@Lee1988; @Raven1988; @Wiencke1989; @Dieck1993; @Connell2010; @Bartsch2012; @Mabin2013], light [@Lee1988; @Altamirano2004; @Wernberg2008; @Arkema2009; @Tait2014], wave motion [@Graham1997a; @Kawamata2001; @Wernberg2005; @FowlerWalker2006; @pedersen2012a] and nutrients [@Jackson1977; @AndrewsHutchings1980; @Field1980b; @Probyn1985; @Guimaraes1996; @Edwards2006]. At shallow depths (2–3m), the canopy tends to reach the surface; however, the structure changes to a mixture of sub-canopy and canopy at greater depths (5–10m), with eventual transition into _L. pallida_ dominated canopy (>10m) [@Rothman2006; @Rothman2010]. Therefore, from the surface it may appear that there is a change in biomass due to reduction in canopy cover with depth; however, due to the presence of sub-canopy kelps at deeper depths the biomass changes insignificantly [@Rothman2006]. Populations of E. maxima dominate the west coast, but more recently have increased in density in False Bay and further east [@Reimers2012], extending their range by 70km along the south coast [@Bolton2012] The range extension of E. maxima has been associated with a possible cooling and increased nutrient supply in the nearshore around the South African south-west coast [@Moloney2005; @Blamey2015b].

@@

@@
